---
title: "Quarto articles"
---

<style type="text/css">
    ul { list-style-type: square; }
</style>


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Once you are ready to create a pkgdown website for your R package, complete the following steps. Note: creating articles can be done before or after initiating pkgdown with `use_pkgdown()`, but I show this after so I can locally build articles to check things like formatting prior to deploying with GitHub actions.

# Install a version of pkgdown that addresses quarto articles rending paths

Before starting, you will need to install [pkgdown](https://github.com/ethanbass/pkgdown) by [ethanbass](https://github.com/ethanbass) using the following code:

```{r}
#| eval: false

devtools::install_github("ethanbass/pkgdown", ref="qmd_articles")
```

*note*: this may have been fix may have been incorporated in rlib's pkgdown on CRAN since I wrote this, so check if pkgdown now supports his patches.

# Initiate pkgdown

Next you will need open the directory for your R package and set up the infastructure of pkgdown website:

```{r}
#| eval: false

use_pkgdown()
```

# Create your article(s)

Now you can create your article(s). There are a couple of ways to do so. 

1. Run `use_article()`. This will create the `vignettes/articles` directory, if it does not yet exist, and add `^vignettes/articles$` to the .Rbuildignore. It will also create a .gitignore file in `vignettes/articles` with the contents `*.html` and `*.R`. It will also create a new .Rmd file in `vignettes/articles`. This route requires some tweaks, as we do not want the .Rmd and rather a .qmd. You will need to:

- Delete the .Rmd file.
- Add a new .qmd file in `vignettes/articles` with no format (it will default to html).
- Add `/.quarto/` to `vignettes/.gitignore`.
- Add `.quarto` to `.gitignore`.

2. You can manually set up the articles using the following checklist:

- Create `vignettes` directory.
- Create `vignettes/articles` directory.
- Create `vignettes/.gitignore` file containing: `*.html`, `*.R`, `/.quarto/` (note do not add ',' in actual file).
- Add a new .qmd file in `vignettes/articles` with no format (it will default to html).
- Add `^vignettes/articles$` to .Rbuildignore.
- Add `/.quarto/` to `vignettes/.gitignore`.
- Add `.quarto` to `.gitignore`.

Add content to your article. I recommend running `pkgdown::build_articles()` to check formatting before deploying with GitHub actions.

# Set up site locally

Once you are satisfied with your article, run `pkgdown::build_site()`. This will give you the opportunity to set up formatting for the whole website using the `_pkgdown.yml` or custom css formatting. You should also add a README and determine your landing page.

# Create GitHub actions to deploy site

When you are ready to deploy your site with GitHub actions (GHA), you need to first set up the page.

```{r}
#| eval: false

pkgdown::use_pkgdown_github_pages()
```

You will then need to change the defualt workflow (`.github/workflows/pkgdown.yaml`) to ensure the correct version of pkgdown is used to build the site through GHA. Specifically, like 37 should be changed from `extra-packages: any::pkgdown, local::.` to `extra-packages: github::ethanbass/pkgdown@qmd_articles, local::.`

Add, commit, and push your changes to GitHub and watch as your site is built!


